---
title: "Analyzing IMDb Scores and Ratings for Netflix Titles"
author: "Maged Saeed Abdo Mostafa Kharshom"

output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    theme: cerulean

---

**Matriculation number:** `30496A`

---

# Introduction

### Objective
The primary aim of this project is to conduct a comprehensive analysis of IMDb scores and ratings, age certification ratings, and the evolution of content quality over time for TV shows and movies on Netflix. This multifaceted approach serves several key objectives.

### Analysis of IMDb Scores
- **Goal**: To identify prevailing trends and patterns in audience preferences.
- **Importance**: 
    - Provides valuable insights for content creators and producers on investment-worthy shows or movies.
    - Sheds light on factors contributing to a show’s or movie’s success on Netflix.

### Examination of Age Certification Ratings
- **Goal**: To understand the target audience for different TV shows and movies on Netflix.
- **Relevance**: 
    - Offers critical information for advertisers targeting specific demographic groups.
    - Serves as a resource for parents and guardians in selecting appropriate content for children.

### Evolution of Content Quality Over Time
- **Goal**: To explore how the quality of content on Netflix has evolved over the years.
- **Approach**: Comparing IMDb scores and votes across different release years.
- **Expected Insights**:
    - Potential shifts in audience preferences and industry standards.
    - Influences on viewers' perceptions and opinions over time.

---

# Data used


#### Source of Data
The dataset used in this analysis was obtained from [Back 2 Viz Basics](https://www.kaggle.com/datasets/thedevastator/netflix-imdb-scores/data), which provides comprehensive information about various titles available on the popular streaming platform. The dataset includes details such as the title's name, its type (whether it is a TV show or a movie), a brief description of the content, the year it was released, age certification rating, runtime (for TV shows: length of episodes; for movies: duration), IMDb score, and IMDb votes. The dataset is publicly available and was downloaded for the purpose of this project.

#### Structure of the Dataset
-**The dataset consists of [5283] rows and [11] columns.** 

The dataset includes the following variables:

- **index**: An integer representing the index of each record.
- **id**: A character identifier for each record.
- **title**: A character representing the title of the movie or TV show.
- **type**: A character indicating whether it’s a “MOVIE” or a “SHOW.”
- **description**: A character containing a brief description of the movie or TV show.
- **release_year**: An integer indicating the year of release.
- **age_certification**: A character specifying the age certification (e.g., “R,” “PG”).
- **runtime**: An integer representing the duration in minutes.
- **imdb_id**: A character containing the IMDb ID for the movie or TV show.
- **imdb_score**: A numeric value representing the IMDb score.
- **imdb_votes**: A numeric value indicating the number of IMDb votes.

The data was provided in CSV format and was imported into R for analysis. Prior to analysis, some data cleaning and preprocessing steps were performed to handle missing values and ensure data consistency.

---

# Methodology

## Data Collection and Preparation


#### Load Necessary Libraries
```{r  warning=FALSE, message=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(plotly)
library(htmlwidgets)
library(knitr)

```
#### Load the new cleaned dataset

```{r}

netflix_data <- read.csv("D:/study/coding for data science/Netflix_Imdb_Project/Netflix_IMDB_Scores/Data/processed_data/cleaned_netflix_data.csv")

```

#### Display the first few rows
```{r}
head(netflix_data)

```
#### Summary statistics
```{r}
knitr::kable(summary(netflix_data))


```

#### Information about columns and data types

```{r}
str(netflix_data)


```

#### Check for missing values
```{r}
knitr::kable(colSums(is.na(netflix_data)))
```

---

# Results


## Analysis of IMDb Scores 

`fig_1a` - **Histogram of IMDb Scores:** This visualization separates the IMDb scores by content type (movies or shows) and displays their distribution. It is a direct approach to analyzing audience rating trends and can show if certain scores are more common for either type of content.


```{r}
fig_1a <- ggplot(netflix_data, aes(x = imdb_score, fill = type)) +
  geom_histogram(binwidth = 0.5, position = "dodge", alpha = 0.7) +
  labs(title = "Distribution of IMDb Scores on Netflix",
       x = "IMDb Scores",
       y = "Number of Titles") +
  theme_minimal()
fig_1a

```

#### Distribution of Types

`fig_1b` - **Pie Chart of Content Types:** This chart gives a quick visual representation of the proportion of movies to TV shows in the dataset, although it doesn't directly relate to IMDb scores.

```{r}
type_counts <- table(netflix_data$type)

fig_1b <- plot_ly(
  labels = names(type_counts),
  values = as.vector(type_counts),
  type = "pie"
) %>%
  layout(title = "Distribution of Types")
fig_1b

```

#### IMDb Scores vs. Runtime

`fig_1c` - **Scatter Plot of IMDb Scores vs. Runtime:** This plot can help to identify if there's a relationship between the length of content and its IMDb rating, which can be a factor in content investment decisions.

```{r}


fig_1c <- ggplot(netflix_data, aes(x = runtime, y = imdb_score)) +
  geom_point() +
  labs(title = "IMDb Scores vs. Runtime",
       x = "Runtime",
       y = "IMDb Score") +
  theme_minimal()
fig_1c

```


#### Average IMDb Score by Type

`fig_1d` - **Average IMDb Score by Type Bar Chart:** This shows whether movies or TV shows have higher average ratings, which is directly useful for content creators and producers.

```{r}
fig_1d <- netflix_data %>%
  group_by(type) %>%
  summarise(average_score = mean(imdb_score, na.rm = TRUE)) %>%
  ggplot(aes(x = type, y = average_score, fill = type)) +
  geom_bar(stat = "identity") +
  labs(title = "Average IMDb Score by Type", x = "Type", y = "Average IMDb Score")
fig_1d
```

#### Boxplot of IMDb Scores by Type

`fig_1e` - **Boxplot of IMDb Scores by Type:** This provides insights into the spread and distribution of ratings for movies and shows, including median scores and outliers.

```{r}
fig_1e <- plot_ly(netflix_data, x = ~type, y = ~imdb_score, type = 'box'
) %>%
  layout(title = 'Boxplot of IMDb Scores by Type',
         xaxis = list(title = 'Type'),
         yaxis = list(title = 'IMDb Score'))
fig_1e



```


#### Top Average IMDb Scores 

`fig_1f` - **Treemap of Top IMDb Scores:** This highlights specific titles with scores and votes around a chosen threshold, which can show which high-rated titles are in this score range, but it may be somewhat arbitrary depending on the choice of threshold and margin.

```{r}

threshold_score <- 7.1
margin_of_error <- 0.3 


suppressWarnings({
  netflix_data$imdb_votes <- as.numeric(as.character(netflix_data$imdb_votes))
  netflix_data$imdb_score <- as.numeric(as.character(netflix_data$imdb_score))
})

top_scores <- netflix_data %>%
  filter(between(imdb_score, threshold_score - margin_of_error, threshold_score + margin_of_error)) %>%
  arrange(desc(imdb_votes), desc(imdb_score)) %>%
  head(16) %>%
  mutate(hover_text = paste("Year:", release_year, 
                            "<br>Type:", type, 
                            "<br>IMDb Score:", imdb_score, 
                            "<br>Votes:", format(imdb_votes, big.mark = ",")))

# Create the treemap plot
fig_1f <- plot_ly(
  data = top_scores,
  labels = ~title,
  parents = rep("", nrow(top_scores)),
  type = 'treemap',
  hoverinfo = 'text',
  hovertemplate = '%{customdata}<extra></extra>',
  customdata = ~hover_text,
  textfont = list(size = 16)
) %>%
  layout(
    title = "Top Average IMDb Scores by Vote Count within Score Range",
    margin = list(l = 10, r = 10, b = 10, t = 40)
  )

fig_1f



```

## Examination of Age Certification Ratings

#### Target Audience Analysis by Age Certification and Content Type

`fig_2a` - **Target Audience Analysis:**This visualization provides a comparative analysis of the number of Netflix titles across different age certifications, separated by content type (Movie or Show). By utilizing a bar chart, the graph distinctly illustrates the count of titles for each age certification category, with a clear differentiation between movies and shows.

```{r}
audience_target_analysis <- netflix_data %>%
  group_by(age_certification, type) %>%
  summarise(count = n(), .groups = 'drop') %>%
  arrange(desc(count))

# Visualize the number of titles targeting each age certification, split by content type
fig_2a <-  ggplot(audience_target_analysis, aes(x = age_certification, y = count, fill = type)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Target Audience Analysis by Age Certification and Content Type",
       x = "Age Certification", y = "Number of Titles",
       fill = "Content Type") +
  theme_minimal()
fig_2a


```

#### Ratings by Age Certification

`fig_2b` - **IMDb Scores by Age Certification:** The fig_2b visualization presents a detailed analysis of IMDb scores categorized by age certification for the content available on Netflix. This is achieved through a series of boxplots, each representing a different age certification category.

```{r}
fig_2b <- ggplot(netflix_data, aes(x = age_certification, y = imdb_score, fill = age_certification)) +
  geom_boxplot() +
  labs(title = "IMDb Scores by Age Certification", x = "Age Certification", y = "IMDb Score")
fig_2b


```

## Evolution of Content Quality Over Time


#### Trends Over Time in Average IMDb Scores:

1.	`fig_3a` - **Trends of IMDb Scores Over Years:** This visualization shows the average IMDb score for titles released each year. By plotting a line graph, you can see trends in how the average quality of content, as rated by IMDb users, has changed over time.



```{r}
fig_3a <- netflix_data %>%
  group_by(release_year) %>%
  summarise(
    average_score = mean(imdb_score, na.rm = TRUE),
    se = sd(imdb_score, na.rm = TRUE) / sqrt(n())  # Standard error
  ) %>%
  ggplot(aes(x = release_year, y = average_score)) +
  geom_ribbon(aes(ymin = average_score - se, ymax = average_score + se), alpha = 0.2) +
  geom_line() +
  labs(title = "Trends of IMDb Scores Over Years", x = "Release Year", y = "Average IMDb Score") +
  theme_minimal()
fig_3a


```

#### Relationship Between Scores and Votes:

`fig_3b` - **IMDb Scores vs. IMDb Votes Scatter Plot:** This scatter plot compares the IMDb scores with the number of votes they've received, which might indicate the popularity and audience engagement with titles of different quality. Using a logarithmic scale for votes takes into account the wide range of vote counts.


```{r}
fig_3b <- plot_ly(netflix_data, x = ~imdb_votes, y = ~imdb_score, type = 'scatter', mode = 'markers',
                  marker = list(size = 5, opacity = 0.5)) %>%
  layout(title = 'IMDb Scores vs. IMDb Votes',
         xaxis = list(title = 'IMDb Votes', type = "log"), # Using log scale for X-axis
         yaxis = list(title = 'IMDb Score'))
fig_3b
```

#### IMDB Scores Over the Years by Type

`fig_3c` - **IMDb Scores Over the Years by Type:** By grouping the data by content type (movies or shows) and plotting IMDb scores over time, you can compare how movies and shows might differ in terms of quality trends. This could reveal if there are distinct patterns in audience perception based on the type of content.



```{r}
fig_3c <- ggplot(netflix_data, aes(x = release_year, y = imdb_score, group = type, color = type)) +
  geom_line(aes(group = interaction(release_year, type)), alpha = 0.3) +  # Use interaction for grouping in geom_line
  geom_smooth(se = FALSE, method = "loess" , formula = y ~ x) +  # Add LOESS smoothed trend line
  labs(title = "IMDb Scores Over the Years by Type", x = "Release Year", y = "IMDb Score") +
  theme_minimal() 
fig_3c



```

---

# Conclusion

In our analysis of Netflix's content library, we focused on evaluating IMDb scores and ratings, age certification distributions, and the evolution of content quality over time. Using a detailed dataset, we revealed key trends in audience preferences, essential for content creators and strategists.

Our investigation highlighted how different content types and age certifications are received by diverse audiences. For instance, the age certification analysis provided insights into Netflix's target demographics, offering valuable information for advertisers and content creators. The trend analysis of IMDb scores over the years indicated shifts in content quality and audience expectations, reflecting evolving industry standards and viewer tastes.

Overall, our study presents a snapshot of Netflix's dynamic content strategy, offering critical insights for stakeholders in the digital entertainment industry. As the platform continues to expand and diversify, such data-driven analyses are crucial in understanding and adapting to the changing landscape of viewer preferences.


